---
title: "RSS_V1"
author: "Yemalin Godonou"
date: "August 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This document will attemp to replicate one of the RSS model in R


##this section is the *the data upload and metadata node*  in the model flow
I grab the report from the metadata and the separate thevariables into two files:

* *factors.txt* : for the binary and nominal variables
* *numeric.txt* : for the numerical variables

```{r load the data in R, eval = TRUE}
rm(list = ls())

# get and set the working directory
getwd()
setwd('C:\\Users\\god9690\\Documents\\BigR\\RSS') #use \\ or /  to separate folder

# load the data in R
## using colclass to load decrease the load time and it is recommended for the large data
# load the factors and numerical and character file in the R

fac_var<-readLines('variables/factors.txt')
num_var <-readLines('variables/numeric.txt')
char_var <-readLines('variables/character.txt')

## get the colum number 
char <-which(colnames(rss)%in%c(fac_var,char_var ))
# create a function to transform the variables into classes
class.var <- function(type.v, dat){
  a<- rep(type.v, length(dat))
  names(a)<-dat
  a
  
}
as.vec
#  factor variables
fac<-class.var("factor", dat = fac_var)

# numeric variables
num <-class.var("numeric", dat = num_var)
char <- class.var("character", dat = char_var)
# create the valueor the colclasses
var <- c(fac, char, num)


# load the data in R and the time to load

rss <- read.csv('APS_nochi.csv', header = TRUE, nrows =10 ,na.strings = c("NA",''))

system.time(rss <- read.csv('APS_nochi.csv', header = TRUE,colClasses = var, na.strings = c("NA",'')))

## time to upload the data:
#   user  system elapsed 
 # 216.58    6.45  224.86 
#ptm<-proc.time()
 
dollar <- function(x){
  b<- gsub('$', '',x)
  b
}



rss[,'PVAM'] = sapply(rss[,'PVAM'], dollar)
rss[,'PVAM'] <- as.numeric(rss[,'PVAM'] )

#elap<- proc.time() -ptm


```

The data has **`r nrow(rss.data)`**  and **`r ncol(rss.data)`** columns It took **`r elap[3]`** to  upload the data in R.

```{r, load_DATA, eval=FALSE}
## structure of the file loaded
str(rss.data)
dimnames(rss.data)

```



```{r change_variable_type, echo=TRUE, eval =FALSE}
# make the policy number as rownames:
rownames(rss)<-rss$POLICY_NUMBER



#rss.data[, (colnames(rss.data) %in% fac_var)]<- as.factor(rss.data[, (colnames(rss.data) %in% fac_var)])


```

## the drop node and the filter nodes

``` {r drop_var, eval =FALSE}
# import the file with the drop variables:
drop<-readLines('variables/drop.txt')
rss.drop <- rss[, !(c(colnames(rss) %in% drop))]


```
## filter nodes can be done with dplyr package or with the base R

``` {r dplyr_method, eval =FALSE }
library(dplyr)
rss.filter<- rss.drop %>%
  filter(AGE <=60, 
         #HEIGHT<=90, HEIGHT>=42,
         #BMI>=14, BMI<=60,
         UNDERWRITE_AMT <=2000000,
         AAAPS ==1,
         CHI ==1,
         SRCE_FILE_YEAR_NUM %in% c("2005", "2006", "2007", "2008", "2009", "2010" ,"2011" ,"2012", "2013", "2014")
         ) 
#          )
##FULLUW
# NT2TO             NT2OT are two varibles probabily indicator
#aviation    avocation   driving     mib         CHI
# CHI_AviaInd CHI_DrivingVioInd CHI_OthArrestInd CHI_SmokingInd CHI_DrugsInd CHI_TreatmentInd CHI_TobaccoMedsInd
# CHI_DrugAlcCategory CHI_AllergyImp CHI_AsthmImp CHI_SleepApnImp CHI_AnxietyStressImp CHI_DpressionImp CHI_PanAttckImp CHI_BldStoolImp
# CHI_EsoRefHrtImp CHI_NerBowImp CHI_StmGIPrbImp CHI_ChestPain CHI_CystTumorLumpImp CHI_PolypImp CHI_Diabetes CHI_HeadacheImp CHI_SeizureImp
#
#CHI_HighBP  CHI_KidneyStonesImp CHI_MolesNevi CHI_SkinCancer CHI_PregInferImp CHI_RefractSurg CHI_SurgeryHospImp
# take a closer look at the character variables


## rss remove the AAPS and the CHI Column
#remove some variable
rem <-c("CHI","AAAPS", "POLICY_NUMBER", 'SRCE_FILE_YEAR_NUM','FCTI','FCT3','DEBI','FCT5')
rss.work <- rss.filter[, !(c(colnames(rss.filter) %in% rem))]
```

## data partition node

```{r data_partition }
library(caret)
set.seed(998)
inTraining <- createDataPartition(rss.work$PM_IND, 
                                  p = .70, 
                                  list = FALSE)

training <- rss.work[ inTraining,]
testing  <- rss.work[-inTraining,]

```

### data transofmration node
We will be using 
#### YeoJOnson tranformation and scaling of the data
####  use newZero transformation

* *create dummy variables for the categorical variables* 
* *create binning to allowfor numeric variables* 

``` {r trasform_V1 }
# create dummy varaible we can use cut or use the packages binr



# create the data 2 preprocess .  will be can be use for tree based also
set(1234)
#The Yeo-Johnson transformation is similar to the Box-Cox model but can accommodate predictors with zero and/or negative values (while the #predictors values for the Box-Cox transformation must be strictly positive.) 
pre.process<- preProcess(training[,!(colnames(training) %in% c('PM_IND'))], method = c("center", "scale", "YeoJohnson", "nzv"))
train.process <- predict(pre.process,training[,!(colnames(training) %in% c('PM_IND'))])
test.process <- predict(pre.process,testing[,!(colnames(testing) %in% c('PM_IND'))])

# add the outcome variables:
train.process$PM_IND<- training$PM_IND
test.process$PM_IND <-  test.process, testing$PM_IND

```


``` {r variable_selection }
# Using filter acording to the http://topepo.github.io/caret/filters.html site
library(doMC)
filterCtrl<- sbfControl(
            functions = treebagSBF,
            method = 'repeatedcv', repeats = 5)

system.time(bagFilter <-sbf(PM_IND ~ .,
                data=train.process,
                sbfControl = filterCtrl
                )
)

## applied the variable selection 
train.filter <- predict(bagFilter, head(train.process )) ####issue with running the result


## select the 
```

```{r randomForest }
library(randomForest)

# tune for the optimal numbers of variables to try to spliting on each node

#bestmtry <- tuneRF(training[,!(colnames(training) %in% ('PM_IND'))], training$PM_IND, 5,ntreeTry = 100, 
#                   stepFactor = 1.5 , improve = 0.01, trace =TRUE, PLOT=TRUE,doBest = FALSE
#                   )
set.seed(155)
ctrl <- trainControl(method = 'repeatedcv',
                     repeats = 3,
                     number =5,
                     classProbs =TRUE,
                     summaryFunction = twoClassSummary
                     )

rss.rf<-train(PM_IND ~ .,data=training,
              method="rf",
              trControl= ctrl,
              prox=TRUE,
              metric ='ROC',
              allowParallel=TRUE)

rss.ranfor<- randomForest(PM_IND ~ .,
                          data=training,
                          mtry=10, 
                          na.action = na.roughfix,
                          ntree =500,
                          keep.forest =TRUE,
                          importance=TRUE)

```
